R Notebook

### 2018-02-01 paper/doc/structure.tex
new

@@

The outline of this chapter is as follows. In Section \ref{ch_generativeProcess} we describe the generative process of the Structural Topic Model (STM).

@@

@@

The STM maximizes the posterior likelihood that the observed data were generated by the above data-generating process using an iterative approximation-based variational expectation-maximization algorithm\footnote{A technical description of this maximization process can be found in \citet{roberts_model_2016}} available in R's stm package \citep{roberts_stm:_2016}. Regularizing prior distributions are used for $\gamma$, $\kappa$ and $\Sigma$. Topic models are usually imprecise as the function we are optimizing has multiple modes, such that the model results can be sensitive to the starting values. Since an ex ante valuation of a model is hardly possible, we compute a variety of different models and compare their posterior probability. This enables us to check how substantively different results are for different model solution \citep{roberts_navigating_2016}.

them with respect to the research problem \citep{gentzkow_text_2017}. To evaluate the accuracy of a model its perplexity can be computed. Similar to cross-validation (VC), different model specifications are compared to how well they do in predicting words within a document \citep{asuncion_smoothing_2012}, \citep{wallach_evaluation_2009}. However, \citet{chang_reading_2009} have proven with large user studies that optimal held out likelihood does not correspond to human perception of semantic coherence of topics. \cite{mimno_optimizing_2011} introduced a measure of coherence of a topic by observing co-occurences of the top $N$ terms of each topic on a document. Another strategy for selecting an appropriate model is to check for residual overdispersion of the variance of the multinomial within the data generating process of the model \citep{taddy_estimation_2012}. Other approaches can be found in \citet{airoldi_reconceptualizing_2010} and \citet{teh_hierarchical_2006}. However, cross-checking some subset of assigned topic distributions to evaluate whether the estimates align with the concept of interest \citep{gentzkow_text_2017}. We apply these manual audits together with numeric optimization based on the topic coherence measure suggested by \citet{mimno_optimizing_2011}.

@@

% ----------------------
% Model Results
% ----------------------
\section{Empirical Evaluation}

\todo{Inspect model with ownership covariates}
\todo{Inspect model with different topic-numbers}

Inference of mixed-membership models, such as the one applied in this paper, has been a thread of research in applied statistics in the past few years \citep{blei_latent_2003} \citep{erosheva_mixed-membership_2004} \citep{braun_variational_2010}. As the posterior can be highly dependent on the model definition, using multiple starting points and clever initialization are successful practices. The takeaway from empirical evaluations of the inference task of mixed-membership models is that the bayesian posterior as good frequentist properties \citep{wallach_evaluation_2009}. \citet{roberts_model_2016} use empirical evaluation, including a frequentist coverage analysis and compare the performance of their STM to a variety of alternatives in \citet{roberts_navigating_2016}.

From the STM we get two matrices of parameters: $\beta$ is a row-normalized $K$-by-$V$ matrix of topic-word distributions, where the entry $\beta_{k,v}$ can be interpreted as the probability of observing the $v$-th word in topic $k$. Similarly $\theta$ is a row-normalized $D$-by-$K$ matrix of topic-word distributions, where the entry $\theta_{d,k}$ can be interpreted as the proportion of words in document $d$ which arise from topic $k$. In this section we will use the $\theta$ and $\beta$ matrices to analyze the model results.

Figure \ref{fig_topic_proportion} displays the topics ordered by their expected frequency across the corpus. To assign a label to each topic, we looked at the most frequent words in that topic and the most representative articles \citep{roberts_model_2016}.

Figure \ref{fig_topic_proportion_diff} plots the expected proportion of topic use in public media minus the expected proportion of topic use in private media. Thus topics more associated with public media appear to the right of zero.

@@

% ----------------------
% Effect of covariates
% ----------------------
\subsection{Differences in topic prevalence}
Since we included news source as a covariate in estimating topical prevalence part within the model, we can estimate the frequency a topic is discussed within a news corpus \citep{roberts_model_2016}. We estimate the conditional expectation of topic prevalence for given document characteristics. More specifically, we estimate a linear model, where the documents are observations, the dependent variable is the posterior probability of a topic and the covariates are the metadata of documents (see equation \ref{eq_1}). The stm-package provides a function that uses the method of composition to incorporates uncertainty in the dependent variable, drawing a set of topic proportions from the variational posterior repeated times and compute the coefficients as the average over all results \citep{roberts_stm:_2016}.

@@

\begin{figure}[H]
	\caption{Topic Correlation}
		\begin{center}
			\includegraphics[width=0.9\textwidth,keepaspectratio]{../figs/topic_correlation.png}
		\end{center}
	\label{fig_topic_correlations}
\end{figure}

@@

@@

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=\textwidth,keepaspectratio]{../figs/jsd.png}
	\caption{Jensen-Shannon (JS) Divergence of word-topic probability}
	\label{fig_jsd}
	\end{center}
\end{figure}

>>>>

\todo{Separate by ownership and explain why article distribution represents real-world.}

@@

@@

The STM maximizes the posterior likelihood that the observed data were generated by the above data-generating process using an iterative approximation-based variational expectation-maximization algorithm\footnote{A technical description of this maximization process can be found in \citet{roberts_model_2016}} available in R's stm package \citep{roberts_stm:_2016}. The process gives us two posterior distribution parameter: (1) $\beta$ is a $K$-by-$V$ matrix (where $K=$ number of topics and $V=$ vocabulary), where the entry $\beta_{kv}$ can be interpreted as the probability of observing the $v$-th word in topic $k$. As we let covariate levels (the news provider) effect the word-distribution of a topic, we will get as many $\beta$ matrices as news provider. (2) $\theta$ is a $D$-by-$V$ matrix of the document-topic distributions, where the entry $\theta_{dk}$ can be interpreted as the proportion of words in document $d$ which arise from topic $k$, or rather as the probability that document $d$ deals about topic $k$. We will use these probability distributions to compare the content of public and private news provider in Section \ref{ch_empirical}.

@@

Inference of mixed-membership models, such as the one applied in this paper, has been a thread of research in applied statistics in the past few years \citep{blei_latent_2003} \citep{erosheva_mixed-membership_2004} \citep{braun_variational_2010}. Topic models are usually imprecise as the function we are optimizing has multiple modes, such that the model results can be sensitive to the starting values. Since an ex ante valuation of a model is hardly possible, we compute a variety of different models and compare their posterior probability. This enables us to check how substantively different results are for different model solution \citep{roberts_navigating_2016}. We then cross-checked some subset of assigned topic distributions to evaluate whether the estimates align with the concept of interest \citep{gentzkow_text_2017}. We apply these manual audits together with numeric optimization based on the topic coherence measure suggested by \citet{mimno_optimizing_2011}.

@@

% ----------------------
% Model Results
% ----------------------
\section{Empirical Evaluation}\label{ch_empirical}

Figure \ref{fig_topic_proportion} displays the topics ordered by their expected frequency across the corpus. To assign a label to each topic, we looked at the most frequent words in that topic and the most representative articles \citep{roberts_model_2016}. Figure \ref{fig_topic_proportion_diff} plots the expected proportion of topic use in public media minus the expected proportion of topic use in private media. Thus topics more associated with public media appear to the right of zero.

@@

% ----------------------
% Effect of covariates
% ----------------------
We use various approaches to investigate the differences between news provider content: (1) First, we use the document-topic probability $\theta$, to estimate the conditional expectation of topic prevalence for given document characteristics. (2) Next, we find all edges between topics where they exhibit a positive correlation of $\theta$ above 0.1 to examine how topics are correlated differently for different news wires, indicating how topics are connected and framed differently in each newswire. Approaches (1) and (2) have been used in \citep{roberts_model_2016} \todo{Noch 1,2,3 studien angeben ,die das gemacht haben.} However, we extend the analysis by (3) calculaing the similarity of the word-topic distribution $\beta$ between the news provider using different distance metrices, to identify which topics are discussed similar and differently.

\subsection{Differences in topic prevalence}
Since we included news source as a covariate in estimating topical prevalence part within the model, we can estimate the frequency a topic is discussed within a news corpus \citep{roberts_model_2016}. We estimate the conditional expectation of topic prevalence for given document characteristics. More specifically, we estimate a linear model, where the documents are observations, the dependent variable is the posterior probability of a topic and the covariates are the metadata of documents (see equation \ref{eq_1}). The stm-package provides a function that uses the method of composition to incorporates uncertainty in the dependent variable, drawing a set of topic proportions from the variational posterior repeated times and compute the coefficients as the average over all results \citep{roberts_stm:_2016}.

@@

\begin{figure}[H]
	\caption{Topic Correlation}
		\begin{center}
		\begin{subfigure}{.7\textwidth}
			\includegraphics[width=\textwidth]{../figs/corrplot1.png}	
			\subcaption{Private news provider}
		\end{subfigure}
		\begin{subfigure}{.7\textwidth}
			\includegraphics[width=\textwidth]{../figs/corrplot2}
			\caption{Public news provider}	
		\end{subfigure}
		\end{center}
	\label{fig_topic_correlations}
\end{figure}

@@

for each news provider we get a respective $\beta$, $K$x$V$ matrix. We then use each row (corresponding to each topic) to calculate the similarity of the word-topic distribution between the news provider using the square root of the Jensen-Shannon (JS) Divergence.

@@

\begin{figure}[H]
	\begin{center}
		\includegraphics[width=0.8\textwidth,keepaspectratio]{../figs/distance}
		\caption{Similarity measures of word-topic probabilities}
	\end{center}
\end{figure}
